MinCall --- MinION end2end deep learning basecaller

### 2017-06-30 paper/main.tex
Adapted bib stuff
Add zero-mean and unit variance preporcessing step
Update

\documentclass[runningheads,a4paper]{llncs}
\usepackage{makeidx}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{footnote}
\usepackage{nccmath}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{amssymb}
% \usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage{acronym}
\usepackage{geometry}
\usepackage[hidelinks]{hyperref}
% \hypersetup{pdftex,colorlinks=true,allcolors=blue
\usepackage{float}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{verbatimbox}

\graphicspath{ {Figures/}{imgs/} }
\begin{document}

@@

The exact error rate metric is unreliable since multiple pipeline tools could be the issue. First the sample is prepared, hopefully, uncontaminated and matching reference genome as close as possible, then sequenced using the MinION device obtaining raw data. Next, our model (or other groups ones) come along, basecall the sequence. To evaluate error rate metric basecalled read is aligned to the reference genome using aligners with their own errors/biases, mostly commonly used BWA-MEM~\cite{li2013aligning} and Graphmap~\cite{sovic2016fast}.
\section{Background}

@@

To eliminate the possibility of overfitting to the known reference, the model is trained and tested on reads from different organisms. Due to limited amount of public available raw nanopore sequence data, ecoli was \emph{divided} into two regions.
Reads were split into train and test portions, depending on which region of ecoli they align.
If read aligns inside first 70\% of the ecoli, it is placed into train set, and if it aligns to the second portion, it is placed into test set. Reads whose alignment overlaps train and test region are not used. Important to note that ecoli genome, and genomes of the majority of other bacteria, is cyclical, so reads with alignments that wrap over edges are also discarded. Total train set consist of over 110 thousand reads.
Overview of the entire learning pipeline is shown in figure \ref{fg:train_pipe}.

@@

\begin{figure}[!h]
    \centering
    \theverbbox
    \caption{Target nucleotide sequence preprocessing. A' is surrogate nucleotide during training. After basecalling surrogate nucleotied are converted to original form.}
    \label{fig:data_preprocessing}
\end{figure}

Due to CTC merged nature during decoding, that is in best matched path adjacent duplicates are merged into one, we preprocess the target nucleotide sequence with surrogate nucleotides, such that each second repeated nucleotide is its surrogate. Example provided in figure~\ref{fig:data_preprocessing}.

@@

% set style - not sure which one is official for papers
% spbasic (springer basic makes sense)
\bibliographystyle{spbasic}
\bibliography{refs}
\appendix
\section{Basecallers}\label{app:basecallers}
Here is currated basecaller list:

>>>>

\documentclass[runningheads,a4paper]{llncs}
\usepackage{makeidx}
\usepackage{makecell}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{footnote}
\usepackage{nccmath}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{amssymb}
% \usepackage{amsthm}
\usepackage[utf8]{inputenc}
\usepackage{acronym}
\usepackage{geometry}
\usepackage[hidelinks]{hyperref}
% \hypersetup{pdftex,colorlinks=true,allcolors=blue
\usepackage{float}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{verbatimbox}
% \usepackage{natbib}
\graphicspath{ {Figures/}{imgs/} }
\begin{document}

@@

The exact error rate metric is unreliable since multiple pipeline tools could be the issue. First the sample is prepared, hopefully, uncontaminated and matching reference genome as close as possible, then sequenced using the MinION device obtaining raw data. Next, our model (or other groups ones) come along, basecall the sequence. To evaluate error rate metric basecalled read is aligned to the reference genome using aligners with their own errors/biases, mostly commonly used BWA-MEM~\cite{li2013aligning} and Graphmap~\cite{sovic2016fast}.

@@

To eliminate the possibility of overfitting to the known reference, the model is trained and tested on reads from different organisms. Due to limited amount of public available raw nanopore sequence data, ecoli was \emph{divided} into two regions.
Reads were split into train and test portions, depending on which region of ecoli they align.
If read aligns inside first 70\% of the ecoli, it is placed into train set, and if it aligns to the second portion, it is placed into test set. Reads whose alignment overlaps train and test region are not used. Important to note that ecoli genome, and genomes of the majority of other bacteria, is cyclical, so reads with alignments that wrap over edges are also discarded. Total train set consist of over 110 thousand reads.
Overview of the entire learning pipeline is shown in figure~\ref{fg:train_pipe}.

@@

\begin{figure}[!h]
    \centering
    \theverbbox
    \caption{Target nucleotide sequence preprocessing}
    \label{fig:data_preprocessing}
\end{figure}

Due to CTC merged nature during decoding, that is in best matched path adjacent duplicates are merged into one, we preprocess the target nucleotide sequence with surrogate nucleotides, such that each second repeated nucleotide is its surrogate. Example provided in figure~\ref{fig:data_preprocessing}. All raw input data were normalized to zero mean and unit variance as it yield superiour perfomance with neural networks.

@@

% set style - not sure which one is official for papers
% spbasic (springer basic makes sense)
% \bibliographystyle{spbasic}
\bibliographystyle{splncs03}
\bibliography{refs}
\appendix
\section{Basecallers}\label{app:basecallers}
Here is currated basecaller list:
